---
title: "**Violence and Peace in the Israel-Palestine Conflict**"
subtitle: "Palestinian Fatalities, Economic Sentiment, and Attitudes of Acceptance Toward Israel's Statehood"
author: "Elisabeth Foster"
institute: "Data Analytics"
date: "May 9, 2023"
output:
  xaringan::moon_reader:
    css: 'metropolis'
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

library(ggplot2)
library(tidyverse)
library(patchwork)
library(readxl)
library(reshape2)
library(vtable)

ab1 = read_csv("abFULL.csv")
ab2 = filter(ab1, !is.na(q101CLEAN),!is.na(q709CLEAN))

sv <- read_excel("Support for Violence and Fatalities Data.xlsx")


```


# Data 
**Sources:**
1. United Nations Office for the Coordination of Humanitarian Affairs 
  - Data: fatality and injury counts for the Israel-Palestine conflict 
  - Years: 2008-2023
  - Level of analysis: individual-level fatalities are counted when they are "the         direct result of confrontations between Palestinians and Israelis in the context      of the occupation and conflict" 

2. Arab Barometer survey data 
  - Data: public opinion surveys conducted across MENA 
  - Years: 2006-2017 (Wave I, II, III, IV)
  - Level of analysis: individual level data with respondents over the age of 18,         located in both rural and urban areas, from a randomly selected household

**Country:** Palestine (West Bank and Gaza)


---
# Methodology
**Analysis 1: Compares annual death total in the West Bank and Gaza.** 
  - Provides background information on levels of violence against Palestinians


**Analysis 2: Uses two survey data questions to evaluate a relationship.**

 - *Q1 Economic Sentiment:* How would you evaluate the current economic situation in your country?
      - Very good
      - Good
      - Bad
      - Very bad 

 - *Q2 Israel's Statehood:* Which of the following statements best expresses your opinion about the problem of Israel and Palestine?
      - The Arab world should accept the existence of Israel as a Jewish state in the       Middle East only when the Palestinians accept it.
      - The Arab world should not accept the existence of Israel as a Jewish state in       the Middle East.


---
# Analysis 1: Palestinian Fatalities 

```{r, echo = FALSE, dpi = 75, fig.retina = 3}
sv1 = 
  sv %>%
  filter(Year > 2007) %>%
  select("Year", "Palestinian Fatalities in WB", "Palestinian Fatalities in G")

sv2 <- melt(sv1, id.vars = "Year")
sv2$value <-as.numeric(sv2$value)

ggplot(sv2,
       aes(x = Year,
           y = value,
           col = variable)) + 
  geom_line() + 
  theme_minimal() + 
  labs( x = "Year",
        y = "Fatalities",
        title = bquote(~bold("Palestinian Fatalities by Territory 2008 to Present Day" )),
        color = "Location") +
  scale_colour_manual(values = c("#fee200", "#008000"),
                      labels = c("West Bank", "Gaza"))

```

---
# Analysis 1 Continued
**Summary Statistics:**
```{r, echo = FALSE}
sv1$`Palestinian Fatalities in WB` <-as.numeric(sv1$`Palestinian Fatalities in WB`)
sv1$`Palestinian Fatalities in G` <-as.numeric(sv1$`Palestinian Fatalities in G`)

sumtable(sv1, 
   title = "Table 1",
   vars = c("Palestinian Fatalities in WB", "Palestinian Fatalities in G"),
   summ = c('mean(x)', 'sd(x)', 'min(x)', 'max(x)'),
   labels = c("Palestinian Fatalities in the West Bank", "Palestinian Fatalities in Gaza"))
```


**Implications and Explanations:**
- Significantly more fatalities in Gaza than in the West Bank
- Spikes in Gaza are a result of three periods of "major hostilities" amongst armed     Palestinian groups in Gaza and Israel
    1. December 2008-January 2009
    2. November 2012
    3. July-August 2014 
- Notable difference in the West Bank and Gaza is important to remember for later       analysis 

---
# Analysis 2: All Palestinian Territories 

```{r, echo = FALSE, dpi = 75, fig.retina = 3}
ggplot(ab2) +
  aes(x=factor(q101CLEAN, level=c('very good', 'good', 'bad', 'very bad')),
        fill = q709CLEAN) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c('#abd7eb', '#f47174')) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Economic Sentiment",
    y = "Percent",
    fill = "Israel's Statehood",
    title = bquote(~bold("Economic Sentiments and Feelings Toward Israel's Statehood")),
    caption = "P-value: 0.1926") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```


---
# Analysis 2: Comparing West Bank and Gaza 

```{r, echo = FALSE, dpi = 75, fig.retina = 3}
abwb = filter(ab1, !is.na(q101CLEAN),!is.na(q709CLEAN),a1=="West Bank")

wbplot = abwb %>%
  ggplot(aes(x=factor(q101CLEAN, level=c('very good', 'good', 'bad', 'very bad')),
      fill = q709CLEAN)) +
  geom_bar(position = "fill", show.legend = FALSE) +
  scale_fill_manual(values = c('#abd7eb', '#f47174')) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Economic Sentiment",
    y = "Percent",
    fill = "Israel's Statehood",
    title = "West Bank"
  ) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45))


abg = filter(ab1, !is.na(q101CLEAN),!is.na(q709CLEAN),a1=="Gaza")

gplot = abg %>%
  ggplot(aes(x=factor(q101CLEAN, level=c('very good', 'good', 'bad', 'very bad')),
      fill = q709CLEAN)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c('#abd7eb', '#f47174')) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Economic Sentiment",
    y = "Percent",
    fill = "Israel's Statehood",
    title = "Gaza"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))

wbplot + gplot +
plot_annotation(title = bquote(~bold("Economic Sentiments and Feelings Toward Israel's Statehood")),
                caption = "P-values: West Bank 0.3487, Gaza 0.0014")

```

---
# Analysis 2 Continued
**Implications and Explanations:**
- There **is not** a clear relationship between economic sentiment and support for Israeli statehood across all of Palestine, however, this differs when splitting the territories.
- There **is** a statistically significant relationship between economic sentiment and support for Israeli statehood in Gaza (p-value 0.0014).
    - As economic sentiment increases, support for Israeli statehood decreases
- Further research could be done to understand why the relationship is different in Gaza. Possible topics to explore include: 
    - The idea that when the economy is doing well, Gazans believe that they are            self-sustaining and thus do not need the help of Israel and/or have the capacity       to actively oppose Israel when their basic needs are met economically.  
